---
title: "README"
output: github_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

# UFC Scraper and Dashboard

This project aims to:

1.  Scrape all available data on UFC fights and fighters:

    1.  Fight and fighter data comes form www.ufcstats.com.
    2.  Betting odds data comes from \_????\_

2.  Display and visualize this data within an R shiny dashboard.

Please note that is work is not intended for commercial gain. It is a task that was set to myself as a means to learn web-scraping and shiny dash-boarding skills. I decided to go with UFC data because I am a fan of MMA and the UFC and I was not satisfied by their official data visualization offerings.

I am new to both web-scraping and dash-boarding, so if anyone happens across any errors or areas in need of improvement, please let me know at [nathan.potgieter56\@gmail.com](mailto:nathan.potgieter56@gmail.com).

The userdefined functions used in this project are located in the functions.R file.

## Web-Scraper

The following packages are used in the web-scraping process:

```{r message=FALSE, warning=FALSE}
if( !require("pacman")) install.packages("pacman")
pacman::p_load(rvest, tidyverse, anytime)
```

Here are the data-sets created:

-   Fighter Data

```{r}
load("data/UFC_Fighters.rda")
as_tibble(df_fighters_full)
```

-   Fight Data

### Fetching Fighter Data

The 'scrape_fighters_raw' function scrapes all the fighter data from a given url.

```{r eval=FALSE, include=TRUE}
source("functions.R")
fighters_raw <- scrape_fighters_raw("http://www.ufcstats.com/statistics/fighters?char=a&page=all")
save("fighters_raw", file = "data/fighters_raw.rda")
```

```{r eval=TRUE, include=FALSE}
load("data/fighters_raw.rda")
print(fighters_raw)
```

Only fighters with last name beginning with A are provided under this URL. Therefore, we have to loop this process over each letter in the alphabet to get the complete dataset.

```{r eval=FALSE, include=TRUE}
source("functions.R")
# This creates a list of 26 urls, 
# once for each letter of the alphabet
url <- "http://www.ufcstats.com/statistics/fighters?char=a&page=all"
lin <- gsub("a&page=all", "", url)
links <- letters %>% map( ~ paste0(lin, ., "&page=all"))

# This scrapes the figher data from each link and 
# reduces each table into a single tibble
df_fighters_raw <- links %>%
    map(scrape_fighters_raw) %>%
    reduce(union) %>% as_tibble()
# save(df_fighters_raw, file = "df_fighters_raw.rda")

```

### Cleaning Fighter Data

```{r}
# Cleaning df_fighters_raw
load("df_fighters_raw.rda")

df_fighters <-
df_fighters_raw %>%
mutate( 
    Name = paste(First_Name, Last_Name),
    .before = First_Name,
    across(all_of(c("Height","Weight", "Reach")), ~gsub("[[:punct:]]", "", .)),
    #across(where(is.character), ~gsub("[[:space:]]", "", .)),
    # across(where(is.character), ~gsub("lbs", "", .)),
    Weight = gsub("lbs", "", Weight),
    # across(where(is.character), ~ gsub("([a-z])([A-Z])", "\\1 \\2", .)),
    across(where(is.character), trimws),
    ht_ft = as.numeric(substring(Height, 1, 1)),
    ht_inch = as.numeric(substring(Height, 2)),
    Height = (0.3048*ht_ft) + (0.0254*ht_inch),
    Reach = 0.00254*as.numeric(Reach)
    ) %>% select(-ht_ft, - ht_inch) 

save(df_fighters, file = "data/df_fighters.rda")
```

## Fetching Additional Data for Each Fighter

By now, if you have been running this code localy, you may have noticed the inclusion of a url column in **df_fighters** and **df_fighters_raw**. This url's was scraped in tandem with the other fighter data. These url's allow us to view additional career stats for each fighter, see for example : <http://www.ufcstats.com/fighter-details/9e8f6c728eb01124>.

In this next step I loop another scraping function (X in functions.R) over each of these url's and then join the resulting output to **df_fighters.**

The additional features include:

-   DOB - Date of birth.
-   SLpm - Significant Strikes Landed per Minute.
-   SApM - Significant Strike Accuracy.
-   SApM - Significant Strikes Absorbed per Minute.
-   Str_Def - Significant Strike Defence (the % of opponents strikes that did not land).
-   TD_Avg - Average Takedowns Landed per 15 minutes.
-   TD_Acc - Takedown Accuracy.
-   TD_Def - Takedown Defense (the % of opponents TD attempts that did not land).
-   Sub_Avg - Average Submissions Attempted per 15 minutes.

```{r eval=FALSE, include=TRUE}
source("functions.R")
load("data/UFC_Fighters.rda")
# Warning this step takes +-30 minuets.
fighter_stats_raw <- UFC_Fighters$Fighter_URL %>% 
    map(scrape_fighter_stats_raw)
# save(fighter_stats_raw, file = "data/df_fighter_stats_raw.rda")
```

Now to format the new data

```{r}
load("data/df_fighter_stats_raw.rda")

fighter_stats <-
fighter_stats_raw %>% reduce(union) %>%
    mutate(DOB = as.Date(anytime(DOB)),
           across(contains("Acc")|contains("Def"), ~ gsub("%|\\.", "", .)),
           across(contains("Acc")|contains("Def"), ~ as.numeric(.)/100),
           across(-c("Fighter_URL", "DOB"), as.numeric))

```

Joining **fighter_stats** with **df_fighters**

```{r}
UFC_Fighters <-
df_fighters %>%
    inner_join( fighter_stats,
               by = c("Fighter_URL" = "Fighter_URL"))
save(UFC_Fighters, file = "data/UFC_Fighters.rda")
```

### Fetching Fight Data

The code below scrapes the Name/date, Location and URL for every UFC event that has occurred to date as well as for the most recent upcoming event. All this information is located [here](http://www.ufcstats.com/statistics/events/completed?page=all).

```{r eval=FALSE, include=TRUE}
# Scraping links to fight event pages
link <- "http://www.ufcstats.com/statistics/events/completed?page=all"
site <- read_html(link) # All the HTML and CSS code from link.

fight_events_dirty <-  site %>%
    html_table(fill = T) %>%
    .[[1]] %>% .[-1,] %>%
    as_tibble() %>%
    mutate(across(where(is.character), noquote),
           across(where(is.character), ~gsub("[[:punct:]]", "", .)),
           across(where(is.character), ~gsub("[[:space:]]", "", .))) %>%
    mutate( url = site %>% # feed `main.page` to the next step
            html_nodes(".b-statistics__table-content") %>% # get the CSS nodes
            html_nodes("a") %>%
            html_attr("href")
    )
# save(fight_events_dirty, file = "data/fight_events_dirty.rda")
```

Cleaning this data:

```{r}
load("data/fight_events_dirty.rda")

month_names <- "January|February|March|April|May|June|July|August|September|October|November|December"

fight_events <-
    fight_events_dirty %>%
    set_names(c("Event", "Location", "Event_URL")) %>%
    mutate(Month = str_extract(Event, month_names),
           Event = str_remove(Event, month_names)) %>%
    separate(Event, into = c("Event", "Date"), sep = -6) %>%
    separate(Date, into = c("Day", "Year"), sep = -4) %>%
    unite(all_of(c("Month", "Day", "Year")), col = "Date", sep = "-") %>%
    mutate(Date = as.Date(anytime(Date)),
           across(all_of(c("Event", "Location")), ~ gsub("([a-z])([A-Z])", "\\1 \\2", .)),
           Event = sub("UFCF", "UFC F", Event),
           Event = gsub('(\\d|\\d\\d|\\d\\d\\d)', ' \\1 ', Event),
           Event = gsub("vs ", " vs ", Event),
           Event = trimws(Event, which = "both") ) %>%
    select(Date, Event, Location, Event_URL)
```

Fetching more detail from **fight_events$url**

```{r eval=FALSE, include=TRUE}
source("functions.R")

fights_raw <- fight_events$Event_URL %>% map(scrape_events)
# save("data\fights_raw.rda")

```

Cleaning the Fight data:

```{r}
pacman::p_load(tidylog)
load("data/fights_raw.rda")

UFC_Fights <- 
    fights_raw %>% 
    #map2(.y = fight_events$Event_URL, ~ mutate(.x, Event_URL = .y) ) %>%
    reduce(union) %>% 
    inner_join(fight_events) %>%  # 100% match
    select(Date, Event, Location, Event_URL, Win_Loss:Time) %>% 
    mutate(
        #Win_Loss = if_else( grepl("win", Win_Loss), "Win   Loss", "nc   nc"),
        #Win_Loss = if(Date >= Sys.Date()) "TBC",
        Win_Loss = if_else( Date >= Sys.Date(), "TBC   TBC", 
                            if_else( grepl("win", Win_Loss), "Win   Loss", "NC   NC") ),
        Fight = Fighter,
        .before = Win_Loss
        ) %>% 
    separate_rows(c("Win_Loss", "Fighter", "Kd", "Str", "Td", "Sub"), sep = "\\s{2,}") %>% 
    separate(Method, into = c("Method", "Method_Detail"), sep = '\\s+', fill = "right", extra = "merge") %>% 
    mutate(
        Fight = sub('\\n', " vs", Fight),
        Fight = gsub('\\s+', " ", Fight),
        across(all_of(c("Kd", "Str", "Td", "Str", "Td", "Sub")), as.integer)
        ) 

save(UFC_Fights, file = "data/UFC_Fights.rda")
```

## Putting it All Together
```{r}
load("data/UFC_Fighters.rda")
load("data/UFC_Fights.rda")

UFC_Data <- UFC_Fights %>% 
    inner_join(UFC_Fighters,  # Fix URL names in each Source
               by = c("Fighter" = "Name")) %>% 
    select(Date:Other,DOB:Sub_Avg, Fighter_URL)

save(UFC_Data, file = "data/UFC_Data")
load("data/UFC_Data.rda")

# UFC_Data %>% filter( grepl("Conor McGregor", Fight, fixed = TRUE) ) %>% View()
# UFC_Fights %>% filter(Fighter == "Justin Gaethje")
# UFC_Fighters %>% filter(Name == "Conor Mc Gregor") # Need to fix missalignment

```


```{r}
load("data/UFC_Fighters.rda")
load("data/UFC_Fights.rda")
load("data/UFC_Data")
UFC_Data %>% filter( grepl("Conor McGregor", Fight, fixed = TRUE) )



n <- 10
order_by <- "SLpM"

UFC_Fighters %>% 
    arrange( desc(order_by) ) %>% 
    slice_max(order_by = SLpM, n = n) %>% 
    ggplot( aes(x = Name, y = order_by, fill = DOB) ) + 
    geom_col() + 
    coord_flip()


```


```{r}
write_csv(UFC_Fighters, file = "data/UFC_Fighters.csv")
```

